Extracting The Power of Dependent Types

### 2021-07-15 paper.tex
use wrapfig to save more space
Kill hanging lines, center fig 1

\usepackage{epsf}

@@

  Most existing programming languages provide little support to formally
  state and prove properties about programs.  Adding such capabilities is far
  from trivial, as it requires significant re-engineering of the existing
  compilers and tools.  This paper proposes a novel technique to write
  correct-by-construction programs in languages without built-in verification
  capabilities, while maintaining the ability to use existing tools.
  This is achieved in three steps.  Firstly, we
  give a shallow embedding of the language (or a subset) into a
  dependently typed language.  Secondly, we write a program
  in that embedding, and we use dependent types to guarantee
  correctness properties of interest.  Thirdly, we
  extract a program written in the original language,
  so that it can be used by the existing compilers and tools.

  Our main insight is that it is possible to
  express all three steps in a single language that supports
  both dependent types and reflection. Essentially,
  this allows us to express a program, its formal properties, and
  a compiler for it hand-in-hand, offering a lot of flexibility to programmers.
  We demonstrate this three-step approach by embedding a subset of
  the PostScript language in Agda, and illustrating it with several short examples.
  Thus we use the power of reflection to bring the benefits of dependent types to languages
  that had to go without them so far.

@@

%\paragraph{Metaprogramming}
There is a large body of
work on metaprogramming facilities in various programming
languages.  \citet{refl-masses} track the origins
of metaprogramming to Smith's work on reflection in Lisp \cite{refl-lisp}.
Some prominent metaprogramming systems include
MetaOcaml \cite{metaocaml}, MetaML \cite{metaml},
reFlect \cite{DBLP:journals/jfp/GrundyMO06},
Template Haskell \cite{sheard2002template},
Racket \cite{plt-tr1}, and various other Lisp/Scheme dialects.
%
However, these systems typically do not support dependent
types, so they are not well suited for our goal of statically
enforcing correctness of embedded programs.

>>>>

\usepackage{epsf}
\usepackage{wrapfig}

@@

  Most existing programming languages provide little support to formally
  state and prove properties about programs.  Ad\-ding such capabilities is far
  from trivial, as it requires significant re-engineering of the existing
  compilers and tools.  This paper proposes a novel technique to write
  correct-by-con\-struc\-tion programs in languages without built-in verification
  capabilities, while maintaining the ability to use existing tools.
  This is achieved in three steps.  Firstly, we
  give a shallow embedding of the language (or a subset) into a
  dependently typed language.  Secondly, we write a program
  in that embedding, and we use dependent types to guarantee
  correctness properties of interest.  Thirdly, we
  extract a program written in the original language,
  so that it can be used by the existing compilers and tools.

  Our main insight is that it is possible to
  express all three steps in a single language that supports
  both dependent typ\-es and reflection. Essentially,
  this allows us to express a program, its formal properties, and
  a compiler for it hand-in-hand, offering a lot of flexibility to programmers.
  We demonstrate this three-step approach by embedding a subset of
  the PostScript language in Agda, and illustrating it with several short examples.
  Thus we use the power of reflection to bring the benefits of dependent types to languages
  that had to go without them so far.

@@

%\paragraph{Metaprogramming}
There is a large body of
work on metaprogramming facilities in various programming
languages.  \citet{refl-masses} track the origins
of metaprogramming to Smith's work on reflection in Lisp \cite{refl-lisp}.
Some prominent metaprogramming systems include
MetaOcaml \cite{metaocaml}, Me\-taML \cite{metaml},
reFlect \cite{DBLP:journals/jfp/GrundyMO06},
Template Haskell \cite{sheard2002template},
Racket \cite{plt-tr1}, and various other Lisp/Scheme dialects.
%
However, these systems typically do not support dependent
types, so they are not well suited for our goal of statically
enforcing correctness of embedded programs.
